A Bioconductor Workflow for Processing and Analysing Spatial Proteomics Data

### 2016-12-02 bioc-workflow.tex
updates based on Claire's comments

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{combined} \hlkwb{<-} \hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
\hlstd{combined}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 6725 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2Y3-3 (6725 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 17:19:06 2016 
##  MSnbase version: 1.21.7
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
\hlstd{combined}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 17:19:06 2016 
## Subset [6725,20][5032,20] Fri Dec  2 17:19:07 2016 
## Removed features with more than 0 NAs: Fri Dec  2 17:19:07 2016 
## Dropped featureData's levels Fri Dec  2 17:19:07 2016 
##  MSnbase version: 1.21.7
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{figure}[!ht]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{prots} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"Q8CG48"}\hlstd{,} \hlstr{"Q8CG47"}\hlstd{,} \hlstr{"Q8K2Z4"}\hlstd{,} \hlstr{"Q8C156"}\hlstd{)}
\hlstd{foi13s} \hlkwb{<-} \hlkwd{FeaturesOfInterest}\hlstd{(}\hlkwc{description} \hlstd{=} \hlstr{"13S consensin proteins"}\hlstd{,}
                             \hlkwc{fnames} \hlstd{= prots,}
                             \hlkwc{object} \hlstd{= hl)}
\hlstd{foi13s}
\end{alltt}
\begin{verbatim}
## Traceable object of class "FeaturesOfInterest"
##  Created on Fri Dec  2 17:19:09 2016 
##  Description:
##   13S consensin proteins
##  4 features of interest:
##    Q8CG48, Q8CG47, Q8K2Z4, Q8C156
\end{verbatim}
\begin{alltt}
\hlkwd{plot2D}\hlstd{(hl)}
\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{cex} \hlstd{=} \hlnum{.6}\hlstd{)}
\hlkwd{highlightOnPlot}\hlstd{(hl, foi13s)}
\end{alltt}
\end{kframe}
\includegraphics[width=.8\textwidth]{figure/foi-1}

@@

% latex table generated in R 3.3.2 by xtable 1.8-2 package
% Fri Dec  2 17:19:09 2016
\begin{table}[ht]
\centering
\begin{tabular}{rrllr}
  \hline
 & Replicate & Tag & Gradient.Fraction & Iodixonal.Density \\ 
  \hline
X127C &   1 & 127C & 8 to 9 (pooled) & 11.00 \\ 
  X127C.1 &   2 & 127C & 7 to 9 (pooled) & 10.00 \\ 
   \hline
\end{tabular}
\caption{Differences in gradient fraction pooling.} 
\label{tab:pdtab}
\end{table}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{hl}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126 X127N ... X131.1 (20 total)
##   varLabels: Replicate Tag Gradient.Fraction Iodixonal.Density
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: uniprot.accession uniprot.id ... pd (27 total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Added markers from  'mrk' marker vector. Fri Dec  2 17:19:08 2016 
## Added markers from  'pdres' marker vector. Fri Dec  2 17:19:09 2016 
##  MSnbase version: 2.1.4
\end{verbatim}
\begin{alltt}
\hlkwd{getMarkers}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{)}
\end{alltt}
\begin{verbatim}
## organelleMarkers
##                          40S Ribosome 
##                                   106 
##                          60S Ribosome 
##                                    95 
## Endoplasmic reticulum/Golgi apparatus 
##                                   393 
##                         Mitochondrion 
##                                   525 
##                           Phenotype 1 
##                                   300 
##                           Phenotype 2 
##                                   253 
##                           Phenotype 3 
##                                   203 
##                           Phenotype 4 
##                                    74 
##                           Phenotype 5 
##                                    91 
##                       Plasma membrane 
##                                   421 
##                            Proteasome 
##                                    92 
##                               unknown 
##                                  2479
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

@@

\end{knitrout}
\caption{Classification results. Colours indicate class membership and
  point size are representative or the classification confidence.}
  \label{fig:plotSVM}
\end{figure}

@@

\begin{figure}[!ht]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{## First remove the markers}
\hlstd{preds} \hlkwb{<-} \hlkwd{unknownMSnSet}\hlstd{(hl)}
\hlcom{## Plot a boxplot of the scores of each organelle}
\hlkwd{par}\hlstd{(}\hlkwc{oma} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{10.5}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{))}
\hlkwd{boxplot}\hlstd{(svm.scores} \hlopt{~} \hlstd{svm,} \hlkwc{data} \hlstd{=} \hlkwd{fData}\hlstd{(preds),}
        \hlkwc{ylab} \hlstd{=} \hlstr{"SVM scores"}\hlstd{,} \hlkwc{las} \hlstd{=} \hlnum{2}\hlstd{)}
\end{alltt}
\end{kframe}
\includegraphics[width=.8\textwidth]{figure/thresholds-1}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{ts} \hlkwb{<-} \hlkwd{orgQuants}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{=} \hlnum{.5}\hlstd{)}
\end{alltt}
\begin{verbatim}
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##               0.4368179               0.3037653               0.3856993 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##               0.6916475               0.5951696               0.4377555 
##    Extracellular matrix                Lysosome           Mitochondrion 
##               0.4317514               0.5919750               0.9527663 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##               0.7947663               0.7096575               0.3269003 
##         Plasma membrane              Proteasome 
##               0.7210996               0.4155492
\end{verbatim}
\begin{alltt}
\hlstd{hl} \hlkwb{<-} \hlkwd{getPredictions}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{= ts)}
\end{alltt}
\begin{verbatim}
## ans
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##                      85                     172                      88 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##                     296                     477                     100 
##    Extracellular matrix                Lysosome           Mitochondrion 
##                      27                     123                     513 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##                     231                     344                      42 
##         Plasma membrane              Proteasome                 unknown 
##                     318                     157                    2059
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

In \texttt{pRoloc} there is functionality for unsupervsied machine
learning methods. In unsupervised learning, the training data consists
of a set of input vectors e.g. protein profiles, ignoring the
information about the class label e.g. localisation, other than for
annotation purposes. The main goal in unsupervised learning is to
uncover groups of similar features within the data, termed
clustering. Ordination methods such as principal components analysis
(PCA) also fall into the category of unsupervised learning methods,
where the data can be projected from a high-dimensional space down to
two or three dimensions for the purpose of visualisation.

As described and demonstrated already above, PCA is a valuable and
powerful method for data visualisation and quality control. Another
application uses hierarchical clustering to summarise the relation
between marker proteins using he \texttt{mrkHClust} function, where
the euclidean distance bewteen average class-specific profiles is used
to produce a dendrogramme describing a simple relationship between the
sub-cellular classes (Figure~\ref{fig:mrkHclust}). The
\texttt{mrkHClust} uses the same defaults as all other function, using
the \texttt{markers} feature variable to define marker proteins. In
the code chunk, we adapt the figure margins to fully display the class
names.

\begin{figure}[!ht]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{par}\hlstd{(}\hlkwc{mar} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{15}\hlstd{,} \hlnum{4}\hlstd{,} \hlnum{1}\hlstd{,} \hlnum{0}\hlstd{))}
\hlkwd{mrkHClust}\hlstd{(hl)}
\end{alltt}
\end{kframe}
\includegraphics[width=.5\textwidth,angle=-90]{figure/hclust-1}

@@

We generally find supervised learning more suited to the task of
protein localisation prediction in which we use high-quality curated
marker proteins to build a classifier, instead of using an entirely
unsupervised approach to look for clusters and then look for
enrichment of organelles and complexes. In the latter we do not make
good use of valuable prior knowledge, and in our experience
unsupervised clustering can be extremely difficult due poor estimates
of the number of clusters that may appear in the data.

>>>>

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{combined} \hlkwb{<-} \hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
\hlstd{combined}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 6725 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2Y3-3 (6725 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 19:09:16 2016 
##  MSnbase version: 1.21.7
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
\hlstd{combined}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 19:09:16 2016 
## Subset [6725,20][5032,20] Fri Dec  2 19:09:16 2016 
## Removed features with more than 0 NAs: Fri Dec  2 19:09:16 2016 
## Dropped featureData's levels Fri Dec  2 19:09:16 2016 
##  MSnbase version: 1.21.7
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{figure}[!ht]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{prots} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"Q8CG48"}\hlstd{,} \hlstr{"Q8CG47"}\hlstd{,} \hlstr{"Q8K2Z4"}\hlstd{,} \hlstr{"Q8C156"}\hlstd{)}
\hlstd{foi13s} \hlkwb{<-} \hlkwd{FeaturesOfInterest}\hlstd{(}\hlkwc{description} \hlstd{=} \hlstr{"13S consensin proteins"}\hlstd{,}
                             \hlkwc{fnames} \hlstd{= prots,}
                             \hlkwc{object} \hlstd{= hl)}
\hlstd{foi13s}
\end{alltt}
\begin{verbatim}
## Traceable object of class "FeaturesOfInterest"
##  Created on Fri Dec  2 19:09:18 2016 
##  Description:
##   13S consensin proteins
##  4 features of interest:
##    Q8CG48, Q8CG47, Q8K2Z4, Q8C156
\end{verbatim}
\begin{alltt}
\hlkwd{plot2D}\hlstd{(hl)}
\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{cex} \hlstd{=} \hlnum{.6}\hlstd{)}
\hlkwd{highlightOnPlot}\hlstd{(hl, foi13s)}
\end{alltt}
\end{kframe}
\includegraphics[width=.8\textwidth]{figure/foi-1}

@@

% latex table generated in R 3.3.2 by xtable 1.8-2 package
% Fri Dec  2 19:09:19 2016
\begin{table}[ht]
\centering
\begin{tabular}{rrllr}
  \hline
 & Replicate & Tag & Gradient.Fraction & Iodixonal.Density \\ 
  \hline
X127C &   1 & 127C & 8 to 9 (pooled) & 11.00 \\ 
  X127C.1 &   2 & 127C & 7 to 9 (pooled) & 10.00 \\ 
   \hline
\end{tabular}
\caption{Differences in gradient fraction pooling.} 
\label{tab:pdtab}
\end{table}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{hl}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126 X127N ... X131.1 (20 total)
##   varLabels: Replicate Tag Gradient.Fraction Iodixonal.Density
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: uniprot.accession uniprot.id ... pd (27 total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Added markers from  'mrk' marker vector. Fri Dec  2 19:09:18 2016 
## Added markers from  'pdres' marker vector. Fri Dec  2 19:09:19 2016 
##  MSnbase version: 2.1.4
\end{verbatim}
\begin{alltt}
\hlkwd{getMarkers}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{)}
\end{alltt}
\begin{verbatim}
## organelleMarkers
##                          40S Ribosome 
##                                   106 
##                          60S Ribosome 
##                                    95 
## Endoplasmic reticulum/Golgi apparatus 
##                                   393 
##                         Mitochondrion 
##                                   525 
##                           Phenotype 1 
##                                   300 
##                           Phenotype 2 
##                                   253 
##                           Phenotype 3 
##                                   203 
##                           Phenotype 4 
##                                    74 
##                           Phenotype 5 
##                                    91 
##                       Plasma membrane 
##                                   421 
##                            Proteasome 
##                                    92 
##                               unknown 
##                                  2479
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

In the code chunk above, we pass the whole \texttt{params} parameter
results and, internally, the first pair that return the highest F1
score are returned (using the \texttt{getParams} function above). It
is advised to always check that these are actually good parameters
and, if necessary, set them explicitly, as shown below.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{hl} \hlkwb{<-} \hlkwd{svmClassification}\hlstd{(hl,} \hlkwc{cost} \hlstd{=} \hlnum{16}\hlstd{,} \hlkwc{sigma} \hlstd{=} \hlnum{0.1}\hlstd{,} \hlkwc{class.weights} \hlstd{= w,} \hlkwc{fcol} \hlstd{=} \hlstr{"markers"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\end{knitrout}
\caption{Classification results. Colours indicate class membership and
  point size are representative of the classification confidence.}
  \label{fig:plotSVM}
\end{figure}

@@

\begin{figure}[!ht]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{## First remove the markers}
\hlstd{preds} \hlkwb{<-} \hlkwd{unknownMSnSet}\hlstd{(hl)}
\hlcom{## Plot a boxplot of the scores of each organelle}
\hlkwd{par}\hlstd{(}\hlkwc{oma} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{10.5}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{))} \hlcom{## sets outer margins}
\hlkwd{boxplot}\hlstd{(svm.scores} \hlopt{~} \hlstd{svm,} \hlkwc{data} \hlstd{=} \hlkwd{fData}\hlstd{(preds),}
        \hlkwc{ylab} \hlstd{=} \hlstr{"SVM scores"}\hlstd{,} \hlkwc{las} \hlstd{=} \hlnum{2}\hlstd{)}
\end{alltt}
\end{kframe}
\includegraphics[width=.8\textwidth]{figure/thresholds-1}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{(ts} \hlkwb{<-} \hlkwd{orgQuants}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{=} \hlnum{.5}\hlstd{))}
\end{alltt}
\begin{verbatim}
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##               0.4359719               0.3027894               0.3807832 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##               0.6994864               0.6018155               0.4232624 
##    Extracellular matrix                Lysosome           Mitochondrion 
##               0.4141776               0.5871408               0.9494709 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##               0.7959998               0.7097865               0.3121881 
##         Plasma membrane              Proteasome 
##               0.7104260               0.4136511
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##               0.4359719               0.3027894               0.3807832 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##               0.6994864               0.6018155               0.4232624 
##    Extracellular matrix                Lysosome           Mitochondrion 
##               0.4141776               0.5871408               0.9494709 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##               0.7959998               0.7097865               0.3121881 
##         Plasma membrane              Proteasome 
##               0.7104260               0.4136511
\end{verbatim}
\begin{alltt}
\hlstd{hl} \hlkwb{<-} \hlkwd{getPredictions}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{= ts)}
\end{alltt}
\begin{verbatim}
## ans
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##                      86                     169                      84 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##                     296                     477                      93 
##    Extracellular matrix                Lysosome           Mitochondrion 
##                      27                     124                     523 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##                     230                     343                      38 
##         Plasma membrane              Proteasome                 unknown 
##                     326                     158                    2058
\end{verbatim}
\end{kframe}
\end{knitrout}

The organelle threshold (\texttt{ts} above) can also be set manually
using an interactive app (see below) or by using a named vector of
thresholds, as shown in the putative example below for 4 organelles.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{(ts} \hlkwb{<-} \hlkwd{setNames}\hlstd{(}\hlkwd{c}\hlstd{(}\hlnum{0.612}\hlstd{,} \hlnum{0.701}\hlstd{,} \hlnum{0.81}\hlstd{,} \hlnum{0.92}\hlstd{),} \hlkwd{c}\hlstd{(}\hlstr{"PM"}\hlstd{,} \hlstr{"Mito"}\hlstd{,} \hlstr{"Golgi"}\hlstd{,} \hlstr{"ER"}\hlstd{)))}
\end{alltt}
\begin{verbatim}
##    PM  Mito Golgi    ER 
## 0.612 0.701 0.810 0.920
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

In \Biocpkg{pRoloc} there is functionality for unsupervsied machine
learning methods. In unsupervised learning, the training data consists
of a set of input vectors e.g. protein profiles, ignoring the
information about the class label e.g. localisation, other than for
annotation purposes. The main goal in unsupervised learning is to
uncover groups of similar features within the data, termed
clustering. Ordination methods such as principal components analysis
(PCA) also fall into the category of unsupervised learning methods,
where the data can be projected from a high-dimensional space down to
two or three dimensions for the purpose of visualisation.

As described and demonstrated already above, PCA is a valuable and
powerful method for data visualisation and quality control. Another
application uses hierarchical clustering to summarise the relation
between marker proteins using he \texttt{mrkHClust} function, where
the euclidean distance between average class-specific profiles is used
to produce a dendrogramme describing a simple relationship between the
sub-cellular classes (Figure~\ref{fig:mrkHclust}). The
\texttt{mrkHClust} uses the same defaults as all other function, using
the \texttt{markers} feature variable to define marker proteins. In
the code chunk, we adapt the figure margins to fully display the class
names.

\begin{figure}[!ht]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{par}\hlstd{(}\hlkwc{mar} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{15}\hlstd{,} \hlnum{4}\hlstd{,} \hlnum{1}\hlstd{,} \hlnum{0}\hlstd{))} \hlcom{## set figure margin}
\hlkwd{mrkHClust}\hlstd{(hl)}
\end{alltt}
\end{kframe}
\includegraphics[width=.5\textwidth,angle=-90]{figure/hclust-1}

@@

We generally find supervised learning more suited to the task of
protein localisation prediction in which we use high-quality curated
marker proteins to build a classifier, instead of using an entirely
unsupervised approach to look for clusters and then look for
enrichment of organelles and complexes. In the latter we do not make
good use of valuable prior knowledge, and in our experience
unsupervised clustering can be extremely difficult due to poor
estimates of the number of clusters that may appear in the data.
